---
title: "HW05bio3d"
author: 'San Luc (PID: A59010657)'
date: "10/26/2021"
output:
  pdf_document: default
  html_document: default
---


# Can you improve this analysis code?
library(bio3d)
s1 <- read.pdb("4AKE") # kinase with drug
s2 <- read.pdb("1AKE") # kinase no drug
s3 <- read.pdb("1E4Y") # kinase with drug

s1.chainA <- trim.pdb(s1, chain="A", elety="CA")
s2.chainA <- trim.pdb(s2, chain="A", elety="CA")
s3.chainA <- trim.pdb(s1, chain="A", elety="CA")

s1.b <- s1.chainA$atom$b
s2.b <- s2.chainA$atom$b
s3.b <- s3.chainA$atom$b

plotb3(s1.b, sse=s1.chainA, typ="l", ylab="Bfactor")
plotb3(s2.b, sse=s2.chainA, typ="l", ylab="Bfactor")
plotb3(s3.b, sse=s3.chainA, typ="l", ylab="Bfactor")

Q1. What type of object is returned from the read.pdb() function?

```{r}
library(bio3d)
?read.pdb
s1 <- read.pdb("4AKE") # kinase with drug
s2 <- read.pdb("1AKE") # kinase no drug
s3 <- read.pdb("1E4Y") # kinase with drug

s1.chainA <- trim.pdb(s1, chain="A", elety="CA")
s2.chainA <- trim.pdb(s2, chain="A", elety="CA")
s3.chainA <- trim.pdb(s3, chain="A", elety="CA")

s1.b <- s1.chainA$atom$b
s2.b <- s2.chainA$atom$b
s3.b <- s3.chainA$atom$b

plotb3(s1.b, sse=s1.chainA, typ="l", ylab="Bfactor", top = FALSE, )
plotb3(s2.b, sse=s2.chainA, typ="l", ylab="Bfactor", top = FALSE, )
plotb3(s3.b, sse=s3.chainA, typ="l", ylab="Bfactor", top = FALSE, )
```

```{r}
hc <- hclust( dist( rbind(s1.b, s2.b, s3.b) ) )
plot(hc)
```




#Q6. How would you generalize the original code above to work with any set of input protein structures?



##scoring rubric:
Total 10 points assigned as follows:

###Documentation:
1 pt - comments on what are the inputs to the function.
1 pt - what the function does and how to use it.
1 pt - what is the output of the function.

#Code:
2 pt - function behaves as desired, producing the correct output and follows assignment
specifications.
2 pt - the code is efficient meaning it uses best practices such as limiting calculation
duplication.
2 pt - code is readable, meaning best practices are used including proper indentation
and whitespace used, relevant variable names, and organized in a logical manner.
1 pt - function code and call executes and is working properly.



#Components that we'll be basing on
s1 <- read.pdb("4AKE")
s1.chainA <- trim.pdb(s1, chain="A", elety="CA")
s1.b <- s1.chainA$atom$b
plotb3(s1.b, sse=s1.chainA, typ="l", ylab="Bfactor")

#input: 
There is technically only one input for this function: the file name.

#Functions' function: 
From there, the read function will pull data from the file, trim it, define the atoms in column B chain A, I think, and plot it onto a scatter plot with lines and height with B-factor. 

#How to use it: 
use the function scatterpdb() and insert the protein's name into the ()

#The output is a scatter plot of the proteins of its atoms and some information about its secondary structures. 



```{r}
scatterpdb <- function(input){
  protein <- read.pdb(input) 
  protein.chainA <-trim.pdb(protein, chain="A", elety="CA")
  protein.b <- protein.chainA$atom$b
  plotb3(protein.b, sse = protein.chainA, typ= "l", ylab = "Bfactor")
}

scatterpdb("4AKE")
scatterpdb("1AKE")
scatterpdb("1E4Y")
```

