---
title: "Machine Learning 1"
author: 'San Luc (PID: A59010657)'
date: "10/22/2021"
output:
  pdf_document: default
  html_document: default
---


Kmeans clustering in R is done with the ‘kmeans()' function. Here we make up some data to test and learn it. 

```{r}
tmp <- c(rnorm(30, 3), rnorm(30, -3))
data <- cbind(x = tmp, y = rev(tmp))
plot(data)
```

Run ‘kmeans()’ set k to 2, nstart to 20. We have to specify how many clusters we want. 

```{r}
km <- kmeans(data, centers = 2, nstart = 20)
km
```

Q. How many points are in each cluster?

```{r}
km$size
```
Q. What ‘component’ of your result object details cluster assignment/membership?

```{r}
km$cluster
```
Q. What ‘component’ of your result object details cluster center?

```{r}
km$centers
```

Q. Plot x colored by the kmeans cluster assignment and add cluster centers as blue points

```{r}
plot(data, col = km$cluster)
points(km$centers, col = 'blue', pch = 15, cex = 2)
```

Hierarchical Clustering
Use the ‘hclust()’ function on the same data as before


```{r}
hc <- hclust(dist(data))
hc
```
hclust has a plot method

```{r}
plot(hc)
```

Since we need to find membership vector, find where to cut the tree by using cutree function and tell it at which height to cut at. (hc = x) 
```{r}
cutree(hc, h = 7)
```

```{r}
grps <- cutree(hc, k = 2)
plot(data, col = grps)

```



Principal component analysis (PCA)

```{r}
url <- "https://tinyurl.com/UK-foods" 
x <- read.csv(url)

```
To count the number of rows and columns

```{r}
dim(x)
```
```{r}
x[,-1]
```

```{r}
url <- "https://tinyurl.com/UK-foods"
x <- read.csv(url, row.names = 1)
x
```

```{r}
barplot(as.matrix(x), col = rainbow(17), beside = TRUE)
```

```{r}
mycols <- rainbow(nrow(x))
pairs(x, col = mycols, pch = 16)
```

PCA can help with this. 

Use base R function for PCA, which is called ‘prcomp()’. This function wants the transpose of our data.

```{r}
pca <- prcomp(t(x))
summary(pca)
```
```{r}
plot(pca)
```
```{r}
attributes(pca)
```

We want to use the  pca$x component for this plot
pca$x component for this plot
```{r}
plot(pca$x[,1:2])
text(pca$x[,1:2], labels = colnames(x))
```


We can also examine the PCA “loadings”, which tell us how much the original variables contribute to each PC. 
```{r}
par(mar=c(10, 3, 0.35, 0))
barplot(pca$rotation[,1], las = 2)
```

One more PCA (ayyy)

```{r}
url2 <- "https://tinyurl.com/expression-CSV"
rna.data <- read.csv(url2, row.names=1)
head(rna.data)
```

```{r}
nrow(rna.data)
```


```{r}
ncol(rna.data)
```
```{r}
colnames(rna.data)
```
```{r}
pca.rna <- prcomp(t(rna.data), scale = TRUE)
summary(pca.rna)
```
```{r}
plot(pca.rna)
```

```{r}
plot(pca.rna$x[,1:2])
text(pca.rna$x[,1:2], labels = colnames(rna.data))
12
```



