---
title: "Class17_Vaccine"
author: 'San Luc (PID: A59010657)'
date: "11/24/2021"
output:
  pdf_document: 
    latex_engine: xelatex
---

First we need to import the vaccine file 

```{r}
vax <- read.csv("covid19vaccinesbyzipcode_test.csv")
head(vax)
```


> How many entries are in this dataset? 

```{r}
nrow(vax)
```

## Let's use the skimr package and skim() function to get a quick overview of the structure of this dataset.

```{r}
skimr::skim(vax)
```

We can also do library() and than call the function directly, but since we are using this function only once, we can do skimr:: to use the skim() function as well. 

We notice that one of the column is a date column, and working with time and dates get annoying quickly. We can use the package called lubridate to help us. 

```{r}
library(lubridate)
```
```{r}
today()
```

let's answer the question how many dates since the first entry and the dataset. 

First we should access the first row of column as of date
```{r}
vax$as_of_date[1]
```
However this kind of date is read as column, so we need to change it into something else. 

```{r}
d <- ymd(vax$as_of_date)
```
We now has changed the date column into something we can use (not character)

```{r}
today()- d[1]
```

To make things simpler, we can also overwrite the data set to the needed format. 

```{r}
vax$as_of_date <- ymd(vax$as_of_date)
```

> Q. When was the dataset updated? what its is the last date in this dataset? how many days since the last update? 

First lets access the last entry in the as of date column 
```{r}
vax$as_of_date[nrow(vax)]
```

then we do the math

```{r}
today() - vax$as_of_date[nrow(vax)]
```

> Q. How many days does the data set span? 

```{r}
vax$as_of_date[nrow(vax)] - vax$as_of_date[1]
```
> Q. How many different ZIP code areas are there? 

What we can do is to use the unique() function and then count them using length()

```{r}
zip <- as.factor(unique(vax$zip_code_tabulation_area))
length(zip)

```
To work with ZIP codes, we can use the package called zipcodeR

To download we this package we use install.packages("zipcodeR", dependencies = TRUE)
```{r}
library(zipcodeR)
```

we can pull census data about ZIP code areas (including median household income etc.)
```{r}
reverse_zipcode(c('92037', "92109") )
```

Let's put this aside and look at something else more interesting. 

Let's focus onto San Diego County. We want to subset the full CA vax dataset down to just San Diego County. 

We could do this with base R. 

```{r}
inds <- vax$county == "San Diego"
nrow(vax[inds,])
```

However, sub-setting can get tedious and complicated quickly when you have multiple things we want to subset by. 

So we gonna use the package called dplyr

```{r}
library(dplyr)
```
let's use the filter() function to do our subsetting from now on as it is more straighforward. 

we want to focus in on San Diego County 

```{r}
sd <- filter(vax, county == "San Diego")
nrow(sd)
```

More complicated subsetting 

```{r}
sd.20 <- filter(vax, county == "San Diego",
                age5_plus_population > 20000)
nrow(sd.20)
```


```{r}
sd.today <- filter(vax, county=="San Diego",
                   as_of_date=="2021-11-23")
```

```{r}
summary((sd.today$percent_of_population_fully_vaccinated))
```

Let's make the histogram of these values 

We can use baseR histogram 
```{r}
hist(sd.today$percent_of_population_fully_vaccinated)
```


let's look at data from La jolla (our school)

 this plot is susceptible to being skewed by ZIP code with small population. 

> Q. what is the population of the 92037 ZIP code area? 

```{r}
sd.lajolla <- filter(sd.today, zip_code_tabulation_area == "92037")
sd.lajolla
```

```{r}
sd.lajolla$age5_plus_population
round(sd.lajolla$percent_of_population_fully_vaccinated, 2)
```

> Q.let's look at the 92122 ZIP code 

```{r}
sd.costeverde <- filter(sd.today, zip_code_tabulation_area == "92122")
sd.costeverde
```

```{r}
sd.costeverde$age5_plus_population
round(sd.costeverde$percent_of_population_fully_vaccinated, 2)
```

Let's create a time course-times series for the 92037 ZIP code 

```{r}
sd.time <- filter(vax, zip_code_tabulation_area == "92037")
sd.time
```

```{r}
library(ggplot2)
```

```{r}
ggplot(sd.time) + 
  aes(x = as_of_date, y = percent_of_population_fully_vaccinated) + 
  geom_point() + 
  geom_line(group=1) + 
  labs(x = "Date", y = "Percent Vaccinated")
```


Let's do for all San Diego county with population as large as 92037

How many lines (ZIP code) do we have? 



```{r}
sd.all <- filter(vax, county == "San Diego",
                 age5_plus_population > 36144)
length(unique(sd.all$zip_code_tabulation_area))
```


```{r}
ggplot(sd.all) + 
  aes(x = as_of_date, y = percent_of_population_fully_vaccinated, group = zip_code_tabulation_area) + 
  geom_line() + 
  labs( x = "Date", y = "Percent Vaccinated")
```


Let's look at big population of california 

```{r}
ca <- filter(vax, age5_plus_population > 36144)
length(unique(ca$zip_code_tabulation_area ))
```
```{r}
ggplot(ca) + 
  aes(x = as_of_date, y = percent_of_population_fully_vaccinated, group = zip_code_tabulation_area) + 
  geom_line(alpha = 0.2) + 
  labs( x = "Date", y = "Percent Vaccinated") + 
  geom_hline(yintercept = 0.6672)
```

What is the mean across the state for these 36k+ population areas? 

```{r}
ca.now <- filter(ca, as_of_date == "2021-11-23")
summary(ca.now$percent_of_population_fully_vaccinated)
```

Now we can use the mean value and come back and plug that back into the ggplot. 








